---
title: "Descriptive Analyses for Air Pollution & Dementia"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=F}
# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

knitr::opts_chunk$set(echo = F, cache=F, cache.comments = F, message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE #, fig.height = 8, fig.width = 8
                      )  

pacman::p_load(kableExtra, tidyverse, table1)    
 
# ggplot settings
theme_set(theme_bw())
theme_update(legend.position = "bottom")

set.seed(1)

image_path <- file.path("..", "manuscript", "images")

# Load files 
source("0_functions.R")
sur_person <- readRDS(file.path("Data", "Output", "sur_person.rda"))
sur <- readRDS(file.path("Data", "Output", "sur.rda")) 

```

# Exposure Quality 

* **results are for 10 year exposure window from the MM model for the first pollutant listed**

* exp_wks_coverage, imp_quality, exact_coverage, and imputed_coverage

**exp_wks_coverage**.  
* Proportion of weeks of desired pollutant exposure with nonmissing location/pollutant info
* for MM and ST regions.
    * MM is slightly smaller for PM2.5 and NO2, as we would expect 
* the imputation quality and exact coverage variables are related to ALL of the known address histories, an Not necessarily the address histories that get used in the exposure models. Meaning that they are more useful/better indicators of the quality of the predictions when exposure coverage is high (e.g., 1.0), and less helpful if exposure coverage is low
* Amanda: "if coverage is low I would expect those rows to be excluded anyway.  If the coverage is low then your exposure for that row will have a really different error structure from exposure windows where coverage is high. In general, for most analyses in e.g. MESA, we exclude any rows where coverage is < 95%." 

**imp_qual10_yr_**.  
* Maximum imputation quality of the locations that fall in the relevant period of interest  
* people without valid addresses in the ST modeing region will have NAs     


```{r}
indicator_variables <- c("exp_wks_coverage10_yr_MM",
                         "exact_coverage10_yr_MM",
                         "imputed_coverage10_yr_MM"#, "imp_qual10_yr_MM"
                         )
lapply(indicator_variables, function(x) {
  t <- filter(sur, 
       pollutant == first(pollutant),
       model == "MM",
       exposure_duration ==10
       ) %>%  
  summary_table(var = x) %>%
  kable(caption = paste("distribution of", x), digits = 3) %>%
  kable_styling()
  
  print(t)
  
})


filter(sur, 
       pollutant == first(pollutant),
       model == "MM",
       exposure_duration ==10
       ) %>%
  summarize(
    n = n(),
    imp0 = mean(imp_qual10_yr_MM==0),
    imp1 = mean(imp_qual10_yr_MM==1),
    imp2 = mean(imp_qual10_yr_MM==2),
    ) %>%  
  kable(caption = "proportion of each imp_qual10_yr_MM value (0-2)", digits = 3) %>%
  kable_styling()
  
  

```


# Visit dates

first and last visit dates. 

```{r}
sur_person$intakedt %>% min() 
sur_person$last_visit %>% max() 

```


# IPW

* distribution of weights in sur_person

* could stratify by covariates

# --> max weight is very high. reduce predictor covariates?

```{r}
sur_person %>%
  drop_na(model_wt) %>%
summary_table(., "model_wt") %>% 
  kable(caption = "distribution of IPW", digits = 2) %>%
  kable_styling()

```


# Exposure 

* e.g., from different models

# --> present differently since primary expsures don't vary over time? just present a table?

```{r}
print("Distribution of air pollution exposure predictions")
sur %>%
  filter(exposure_duration ==10) %>%
  group_by(pollutant, exposure_year, model) %>%
  alt_boxplot(var = "pollutant_prediction", min_q=0.1, max_q=0.9) %>% 
  label_pollutants() %>%
  
  ggplot(aes(x=exposure_year, fill=model, group=interaction(exposure_year, pollutant, model))) + 
  geom_boxplot(aes(ymin = Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), 
               stat = "identity") + 
  facet_wrap(~pollutant, scales="free") + 
  labs(x= "Exposure Assessment Year",
       y = "Pollutant Prediction",
       fill = "Model"
       )
  
  
```

# --> START HERE

# --> exposure table (#s for plot above)

```{r}

```


# --> correlation coefficients of predictions from 2019 MM & 2019 SP?

```{r}


```






# Table 1
n, person-years
# --> could startify by high/low exposure like Rachel did, but have to pick a pollutant? what have others done?

# --> change birth cohort?
# --> round %s to whole #s?

```{r}
# Table1 tutorials: 
## https://cran.r-project.org/web/packages/table1/table1.pdf 
## https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html
sur_person$apoe_available <- !is.na(sur_person$apoe)
  
t1_data <- list(
  labels = list(
    male = "Sex",
    intakeage = "Age at Intake",
    follow_up_years = "Follow-Up Years",
    corrected_anydementia = "All-Cause Dementia",
    ad_nincds = "Alzheimer's Disease Cases",
    birth_cohort = "Birth Cohort",
    race_white = "Race",
    degree = "Degree",
    income_cat = "2,000 Census Tract Income at Baseline",
    apoe = "APOE e4 Allele",
    apoe_available = "APOE Status Available",
    model_wt = "IPW for Modeling Cohort"
    ),
  units = list(
    intakeage = "Years"
     ),
  categoricals = list(
    male = list('1' = "Male",
                '0' = "Female"),
    corrected_anydementia = list('1' = "Dementia",
                                 '0' = "No Dementia"
                                 ),
    ad_nincds = list('1' = "Alzheimer's Disease",
                                 '0' = "No Alzheimer's Disease"
                                 ),
    birth_cohort = list(sort(unique(sur_person$birth_cohort))),
    race_white = list('1' = "White",
                      '0' = "Non-White"
                      ),
    degree = list('0' = "None",
                  '1' = "GED/High School",
                  '3' = "Bachelor's",
                  '4' = "Master's",
                  '5' = "Doctorate",
                  '6' = "Other"
                  ),
    income_cat = list('1' = "<$35,000",
                      '2' = "$35,000 - <$50,000",
                      '3' = "$50,000 - <$75,000",
                      '4' = ">$75,000"
                      ),
    apoe = list('1' = "Carriers",
                '0' = "Non-Carriers"
                ),
    apoe_available = list('TRUE' = "APOE Available",
                          'FALSE' = "APOE Not Available"
                          )
    )
  ) %>%
  t1read(sur_person, .) 

# Main Table 1
t1_data %>%
  table1(~intakeage + follow_up_years + male +corrected_anydementia + ad_nincds + race_white +degree + income_cat + apoe + model_wt + birth_cohort , data=.)


# SI, by APOE STATUS/who is in the mode
t1_data %>%
  table1(~intakeage + follow_up_years + male +corrected_anydementia + ad_nincds + race_white +degree + income_cat + apoe + model_wt + birth_cohort | apoe_available, data=.)



```






# Prediction Maps

# --> or reference ACT-TRAP?

```{r}


```

 

