---
title: "Survival Analysis Results for AP & Dementia"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=F}
# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

knitr::opts_chunk$set(echo = F, cache=F, cache.comments = F, message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE)  

pacman::p_load(kableExtra, 
               tidyverse)    
 
source("functions.R")
output_data_path <- file.path("Data", "Output")

# ggplot settings
theme_set(theme_bw())
theme_update(legend.position = "bottom")

set.seed(1)

image_path <- file.path("..", "manuscript", "images")
#if(!file.exists(output_data_path)) {dir.create(output_data_path)} 
if(!file.exists(file.path(image_path, "Other", "HR Plots"))) {dir.create(file.path(image_path, "Other", "HR Plots"))} 

#modeled units
load(file.path(output_data_path, "trap_model_units.rda"))

# Load files 

hrs0 <- readRDS(file.path("Data", "Output", "hazard_ratios.rda"))

main_models <- str_subset(unique(hrs0$description), "Main" )

hrs <- filter(hrs0, grepl("MM_|SP_|ST_", covariate)) %>%
  mutate( 
    covariate = paste(substr(covariate, 4, str_length(covariate)), substr(covariate, 1, 2), sep = "_"),
    covariate = substr(covariate, 1, str_length(covariate)-3)
    ) %>% 
  rename(pollutant = covariate) %>% 
  label_pollutants() %>%
  mutate(pollutant = gsub(pattern = "PNC (", replacement = paste0("PNC (", pnc_units, " "), x = pollutant, fixed = T), 
         pollutant = gsub(pattern = "BC (", replacement = paste0("BC (", bc_units, " "), x = pollutant, fixed = T), 
         pollutant = gsub(pattern = "NO2 (", replacement = paste0("NO2 (", no2_units, " "), x = pollutant, fixed = T), 
         pollutant = gsub(pattern = "PM2.5 (", replacement = paste0("PM2.5 (", pm25_units, " "), x = pollutant, fixed = T),
         pollutant = gsub(", ", "\n", pollutant),
         outcome = ifelse(grepl("dementia", outcome), "All-Cause Dementia", "Alzheimer's Disease"),
         pollutant_predictors = str_to_upper(pollutant_predictors),
         pollutant_predictors = gsub("UFP_10_42|UFP_10_70|UFP_20_1K|UFP_36_1K", "PNC", pollutant_predictors),
         pollutant_predictors = gsub("PM25", "PM2.5", pollutant_predictors),
         # --> update if add models
         pollutant_predictors = factor(pollutant_predictors, levels = c("PNC", "BC", "NO2", "PM2.5",
                                                                        "PNC + PM2.5", "BC + PM2.5", "NO2 + PM2.5",
                                                                        "PNC + BC + NO2 + PM2.5")),
         
         significant = ifelse((lower95 <1 & upper95<1) | (lower95 >1 & upper95>1), "Significant", "Non-Significant" ),
         
         hr_label = paste0(round(hr,2), " (", round(lower95, 2), "-", round(upper95, 2), ")"),
         hr_label = ifelse(significant == "Significant", paste0("*", hr_label), hr_label),
         description = relevel(factor(description), ref = main_models),
         exposure_duration = factor(exposure_duration, levels = rev(unique(exposure_duration))),
         exposure_model = gsub("SP", "2019 ST", exposure_model),
         exposure_model = gsub("MM", "2019 MM", exposure_model),
         )


  
```


# HR Plots

* most of the significant HRs come from the SP exposure model
* 10 yr PNC from the 2005+ exposure year models are the only significant PNC HRs   

```{r}
# dt = hrs
# description. = "Main Models"
# exposure_duration. = "10"

# fn returns HR plots
plot_hrs <- function(dt, description., exposure_duration.) {
  
  p <- dt %>%
    filter(description == description., 
           exposure_duration == exposure_duration.) %>%  
    
    ggplot(., aes(y=pollutant, col=exposure_model, shape = significant, 
                  x=hr, xmin=lower95, xmax=upper95,)) + 
    facet_grid(rows = vars(pollutant_predictors), cols = vars(outcome), 
               space="free", scales="free_y", switch = "both") + 
    geom_vline(xintercept = 1, alpha=0.5, linetype=2) + 
    geom_pointrange(position = position_dodge(width = 1))+
    #geom_text(aes(label=hr_label, x=Inf, hjust=1), position = position_dodge(width = 1), size=3)+
    labs(title = paste0(description., "\n", exposure_duration., " Yr Exposure" ), 
         x= "HR", y="Pollutant Model", col = "Exposure Model", shape="")
  
  return(p)
}

# plot_hrs(hrs, description. = "Main Models (2005+)", exposure_duration. = "10")
# ggsave(file.path(image_path, "Other", paste0("HR_test.png")), width = 8, height = 10)

```

note units:   
* PNC `r pnc_units`   
* BC `r bc_units`   
* NO2 `r no2_units`
* PM2.5 `r pm25_units`


"2005+" Models subset to individuals not yet diagnosed with dementia at the start of 2005 (vs 1994/5). Exposure models go back to 1995.


some observations   
* don't see much of an effect w/ screened p-trak and a slightly higher effect for NanoScan & DiSCmini b/c of the different particle sizes they capture? 

```{r}
pdf(file.path(image_path, "Other", "HR plots", paste0("HR_plots_", Sys.Date(), ".pdf")), width = 8, height = 10)

lapply(group_split(hrs, description, exposure_duration), function(x) {
  plot_hrs(x, description. = first(x$description), exposure_duration. =  first(x$exposure_duration))
  })

dev.off()

```

# --> save main paper plots

```{r}


```





# HR vs TRAP concentration plots 

optional, but have to rerun models w/ diff pollutant increments to get the 95% CI estimated; or use delta method? e.g., Fig 1 in https://www.sciencedirect.com/science/article/pii/S016041202100430X?via%3Dihub

```{r}

```






