---
title: "DAGs for TRAP and Dementia"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(SESsionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(SESsionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

knitr::opts_chunk$set(echo = F, cache=F, cache.comments = F, message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE #, fig.height = 8, fig.width = 8
                      )  

pacman::p_load(tidyverse, dagitty, ggdag)

image_path <- file.path("..", "manuscript", "images")
if(!dir.exists(file.path(image_path, "SI"))) {dir.create(file.path(image_path, "SI"))}

theme_set(theme_dag())

```

# Tutorials

* https://cran.r-project.org/web/packages/ggdag/vignettes/intro-to-dags.html
* https://evalf20.clasSES.andrewheiss.com/example/dags/ 

# Relationships
 
# --> should there not be an arrow from TRAP --> dementia? 

```{r}

dag_labels <- c("Dementia" = "Dementia",
                             "TRAP" = "TRAP",
                             "BMI" = "BMI",
                             "CVHealth" = "Cardiovascular\nHealth",
                             "Smoking" ="Smoking",
                             "Surv65" = "Survival\nto ≥65yr",
                             "APOE" = "APOE",
                             "Sex" = "Sex",
                             "PhysAct" = "Physical\nActivity",
                             "Year" = "Year",
                             "SES" = "SES"
                             )

full_dag <-dagify(Dementia~TRAP+BMI+CVHealth+Smoking+Surv65+APOE+Sex+PhysAct+Year+SES,
                        TRAP~SES+Year+PhysAct,
                        BMI~TRAP,
                        CVHealth~BMI+TRAP+PhysAct+Sex+Smoking,
                        Smoking~SES,
                        Surv65~Smoking+CVHealth+BMI+TRAP+SES+Year+PhysAct+Sex,
                        PhysAct~Sex,
                        outcome = "Dementia",
                        exposure = "TRAP",
                        labels = dag_labels,
                      coords = list(x =c(Dementia = 13,
                                         TRAP =1,
                                         BMI=2,
                                         CVHealth=2,
                                         Smoking=5,
                                         Surv65=5,
                                         APOE=14,
                                         Sex=9,
                                         PhysAct=9,
                                         Year=12,
                                         SES=12,
                                         Smoking=12
                                         ),
                                    y = c(Dementia = 4,
                                         TRAP =4,
                                         BMI=2,
                                         CVHealth=6,
                                         Smoking=1,
                                         Surv65=7,
                                         APOE=1,
                                         Sex=7,
                                         PhysAct=1,
                                         Year=6,
                                         SES=2,
                                         Smoking=7)))

# ggdag_status(full_dag)

final_dag <- full_dag
node_size <- 23

```

# DAGs

* basic DAG 

```{r}
#simple dag
ggdag_status(final_dag, node_size = node_size-2) #+ guides(color = FALSE) 

ggsave(file.path(image_path, "SI", "dag_simple.png"), width = 10, height = 6)

```

* adjusting for colliders and mediators can introduce bias. Instead, we’ll look at minimally sufficient adjustment sets: sets of covariates that, when adjusted for, block all back-door paths, but include no more or no less than necessary. 

```{r}
#adjustment dag
## minimum 
ggdag_adjustment_set(final_dag, #text=F, use_labels = "label", 
                     shadow=T,
                     #text_size = 2, 
                     node_size = node_size,
                     type = "minimal" #default
                     ) + 
  labs(title = "Minimum adjustment covariates")

ggsave(file.path(image_path, "SI", "dag_adjustment_min.png"), width = 10, height = 6)

```

* maximum adjustment covariates that will still produce unbiased estimates

```{r}
## max
ggdag_adjustment_set(final_dag, #text=F, use_labels = "label", 
                     shadow=T,
                     text_size = 2, 
                     node_size = node_size/2,
                     type = "all"  
                     )+ 
  labs(title = "All possible adjustment covariates")

ggsave(file.path(image_path, "SI", "dag_adjustment_max.png"), width = 10, height = 6)

```

*Variables are d-connected if there are any active paths between them and d-separated if there are no open paths between them 

* TRAP and Dementia are d-connected based on the nodes we included

```{r}
# d-relationship
ggdag_drelationship(final_dag,
                    #text_size = 2,
                    node_size = node_size,
                    #text=F, use_labels = "label",  
                    controlling_for = c("SES", "APOE", "Year", "Sex", "Surv65", "PhysAct"
                                        )) + guides(color = FALSE) 

ggsave(file.path(image_path, "SI", "dag_relationship.png"), width = 10, height = 6)

```

* d-paths between TRAP and dementia

```{r}
ggdag_paths(final_dag)

```

# Other 

```{r, results="hide"}

tidy_dagitty(final_dag) %>% dag_adjustment_sets()

```

