---
title: "DAGs for TRAP and Dementia"
author: "Magali Blanco"
date: "6/18/2022"
output: html_document
---

```{r setup, include=FALSE}
# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

knitr::opts_chunk$set(echo = F, cache=F, cache.comments = F, message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE #, fig.height = 8, fig.width = 8
                      )  

pacman::p_load(tidyverse, dagitty, ggdag)

image_path <- file.path("..", "manuscript", "images")

```

# Notes

* tutorial: https://cran.r-project.org/web/packages/ggdag/vignettes/intro-to-dags.html

# Relationships
 
# --> should there not be an arrow from trap --> dementia? 

```{r}
 
# theme_set(theme_dag())
dag_labels <- c("dementia" = "DEMENTIA",
                             "trap" = "TRAP",
                             "bmi" = "BMI",
                             "cv_health" = "Cardiovascular\nHealth",
                             "smoking" ="Smoking",
                             "sur_65" = "Survival\nto ≥65yr",
                             "apoe" = "APOE",
                             "sex" = "Sex",
                             "phys_activity" = "Physical\nActivity",
                             "year" = "Year",
                             "ses" = "SES"
                             )

sur_dag_full <-dagify(dementia~trap+bmi+cv_health+smoking+sur_65+apoe+sex+phys_activity+year+ses,
                        trap~ses+year+phys_activity,
                        bmi~trap,
                        cv_health~bmi+trap+phys_activity+sex+smoking,
                        smoking~ses,
                        sur_65~smoking+cv_health+bmi+trap+ses+year+phys_activity+sex,
                        phys_activity~sex,
                        outcome = "dementia",
                        exposure = "trap",
                        labels = dag_labels)

sur_dag_simple <-dagify(dementia~trap+sur_65+apoe+sex+year+ses,
                               trap~ses+year,
                               sur_65~trap+ses+year+sex,
                               outcome = "dementia",
                                exposure = "trap",
                               labels = dag_labels) 


final_dag <- sur_dag_full

```

# DAGs

adjusting for colliders and mediators can introduce bias. Instead, we’ll look at minimally sufficient adjustment sets: sets of covariates that, when adjusted for, block all back-door paths, but include no more or no less than necessary. 

```{r}
#simple dag
ggdag(final_dag, text=F, use_labels = "label") + theme_dag()
ggsave(file.path(image_path, "SI", "simple_dag.png"), width = 10, height = 6)

#adjustment dag
ggdag_adjustment_set(final_dag, #text=F, use_labels = "label", 
                     shadow=T,
                     text_size = 2) +
  theme_dag()
ggsave(file.path(image_path, "SI", "adjustment_dag.png"), width = 10, height = 6)


```

*Variables are d-connected if there are any active paths between them and d-separated if there are no open paths between them 

* are paths are active or deactive


# --> why are trap-dementia d-connected? the way we coded this?

```{r}
sur_dag_full2 <-dagify(dementia~bmi+cv_health+smoking+sur_65+apoe+sex+phys_activity+year+ses,
                        trap~ses+year+phys_activity,
                        bmi~trap,
                        cv_health~bmi+trap+phys_activity+sex+smoking,
                        smoking~ses,
                       # --> don't inlcude year? although life expectancy increases over time
                       sur_65~smoking+cv_health+bmi+trap+ses+phys_activity+sex+year,
                        phys_activity~sex,
                        outcome = "dementia",
                        exposure = "trap",
                        labels = dag_labels)


# d-relationship
ggdag_drelationship(sur_dag_full2,  
                    text_size = 2,
                    #text=F, use_labels = "label",  
                    controlling_for = c("ses", "apoe", "year", "sex", "sur_65", "phys_activity"
                                        )) + 
  theme_dag()
ggsave(file.path(image_path, "SI", "relationship_dag.png"), width = 10, height = 6)


```

