---
title: "Time-Dependent Effet of PM2.5 on Dementia"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

# Purpose
* compare the health inference results from the old and latest PM2.5 model  

* better understand if there is a time-dependent effects of PM2.5 exposures on dementia incidence and if so, possible explanations. Could the effect be stronger for earlier years because:   

1. the PM2.5 exposure models have changed  
2. Concentrations and variability was higher?   
3. The PM composition was different

* understand how this may relate to the null results we are seeing for UFP, BC & NO2 on dementia incidence using 2019 mobile monitoring exposure models

## Shaffer et al. 2021  
* reported a HR of dementia incidence per 1 ug/m3 increase in PM2.5 of **1.16 (1.03,1.31)**    

* some differences between the latest analyses compared to hers are:
  * a larger sample size
  * an updated ST exposure model with more data & some slight model fit differences 
  * Rachel used forward stepwise selection, I am using Lasso to develop APOE missingness weights. This should be minor since IPW did not change the results much anyways - weights were very close to 1 for both approaches.
  * Rachel's model adjusted for Census median household income, we are now using nses_z_cx:   
    * dementia(age time axis|apoe) ~ PM2.5(age) + degree + sex + median_household_income + race + 2yr_cal_time

  * the current analytic cohort/dataset is based on the person-years with a coverage threshold > 0.5 for a 10 yr MM predictions since PM2.5 was initially a sensitivity analysis (**I could update this but I don't think it will make much of a difference**)

# Overall Approach

* compare PM2.5 exposure models.  

* look at PM2.5 & NO2 effect stratified by time.     
* compare ST PM2.5 results to the 2019 PM2.5 ST model, which adds exposure measurement error like the 2019 MM exposure models, although PM2.5 and UFPs are not strongly correlated, and their surfaces may have changed differently over time.  

* could look at specitation data in the region to look at how PM2.5 composition has changed over time? 


```{r setup, include=FALSE, ECHO=F}
knitr::opts_chunk$set(echo = F, cache=F, cache.comments = F, message = F, warning = F)

######################################################################
# SETUP
######################################################################
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, lubridate, kableExtra, survival, table1) 

set.seed(1)

# ggplot settings
theme_set(theme_bw())
theme_update(legend.position = "bottom")

######################################################################
# NEW EXPOSURE MODELS 
######################################################################
fp <- file.path("Data", "Output")

#load(file.path(fp, "sur_w.rda"))
sur_w <- readRDS(file.path(fp, "sur_w.rda"))
#load(file.path(fp, "cox_model_fns.rda"))
load(file.path(fp, "cox_model_fns.RData"))

main_exposure_duration <- 10
start_of_bad_yrs <- 2019 
 
sur <- sur_w %>%
  filter(exposure_duration==main_exposure_duration) %>%
  drop_na(ST_pm25)

######################################################################
# COMPARE TO OLD EXPOSURE MODELS
######################################################################

# note: no coverage threshold was used before, but coverage was very high, with <3% person-years having one below 98%
old_sur <- readRDS(file.path("Data", "phd_analysis", "20210203 Survival", "dem.w.rda")) %>%
  select(study_id,  cal_2yr= cal_t, dementia_now, age_start_exposure, age_end_exposure, exposure_year, 
         apoe, male, income_cat, race_white, degree, ST_pm25 = pm25_10yr, model_wt)  %>%
  mutate(dataset = "old",
         exposure_duration = 10)  %>%
  drop_na(ST_pm25) %>%  
  # why do i have to do this?
  distinct()

old_ids <- distinct(old_sur, study_id, exposure_year)

sur_new_old_yrs <- left_join(old_ids, mutate(sur, dataset="new" )) %>%
  select(names(old_sur))

comp <- rbind(old_sur, sur_new_old_yrs) %>%
  # is this b/c new dataset has more missingness given our threshold limites?
  drop_na(ST_pm25)

######################################################################
# COMMON VARIABLES
######################################################################

# year start/end dates based on the cuts in the cohort table below
g1_end <- 2000

g2_start <- 2001
g2_end <- 2006

g3_start <- 2007
g3_end <- 2012

g4_start <- 2013
```

# OLD VS NEW MODELS   

* these analyses use the new data predictions but only for the old Study IDs and years to increase comparability 
* The current TRAP survival analyses use a coverage threshold of 0.5, so the new dataset has slightly fewer person-years than the original, which didn't have a threshold, but <3% of the person-years had a threshold < 98% anyway.


```{r}
comp %>%
  drop_na(ST_pm25) %>%
  group_by(dataset) %>%
  summarize(rows = n()) %>%
  kable() %>%
  kable_styling()
  
```


## exposures


```{r}
print("keeping the same id-years from the old dataset in the new prediction dataset")
comp %>% 
  mutate(exposure_year = factor(exposure_year)) %>%
  ggplot(aes(x=exposure_year, y=ST_pm25, col=dataset)) + 
  geom_boxplot() + 
  labs(y = "Predicted 10 yr Exposure", col = "PM2.5 ST Model")

print("there are some strange patterns in the extremes")

comp %>%
  select(study_id, exposure_year, dataset, ST_pm25) %>%  
  pivot_wider(names_from = dataset, values_from = ST_pm25) %>% 
  mutate(exposure_year = factor(exposure_year)) %>% 
  #why have to do this??
  drop_na(old, new) %>%
  
  ggplot(aes(x=old, y=new, col=exposure_year)) + 
  #facet_wrap(~cal_2yr, scales="free") +
  geom_point(alpha=0.05) + 
  geom_smooth() + 
  geom_abline(slope = 1, intercept = 0) +
  labs(title="10 yr PM2.5 predictions")
  
```

## HR models

* unlike our current survival models, these models adjust for income category (not NSES)  

```{r, results="hide"}

dt <- c("old", "new")

models_old_new_time <- lapply(dt, function(x) {
  
  m1 <- comp %>%
    filter(dataset==x) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = "All Years (1994-2018)")

  m2 <- comp %>%
    filter(dataset==x,
           exposure_year <= g1_end) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0("1994-", g1_end))
  
  m3 <- comp %>%
    filter(dataset==x,
           exposure_year >= g2_start & exposure_year <= g2_end) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0(g2_start, "-", g2_end))
  
  m4 <- comp %>%
    filter(dataset==x,
           exposure_year >= g3_start & exposure_year <=g3_end) %>% 
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0(g3_start, "-", g3_end))
  
  m5 <- comp %>%
    filter(dataset==x,
           exposure_year >= g4_start) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0(g4_start, "-2018"))
  
  #others
  ## drop last years where interval censoring might be an issue (incidence is artificially high)
  m6 <- comp %>%
    filter(dataset==x,
           exposure_year <= 2017) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0("Stop early: 1994-", 2017))
  
  m7 <- comp %>%
    filter(dataset==x,
           exposure_year < g4_start) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0("Stop early: 1994-", g4_start-1))
  
  m8 <- comp %>%
    filter(dataset==x,
           exposure_year >= g4_start & exposure_year <= 2017) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0("start 2013: ", g4_start, "-", 2017))
  
  m9 <- comp %>%
    filter(dataset==x,
           exposure_year >= g4_start & exposure_year <= start_of_bad_yrs-3) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
            other_predictors = c("cal_2yr", "degree", "income_cat", "male", "race_white")) %>%
    mutate(description = paste0("start 2013: ", g4_start, "-", start_of_bad_yrs-3))
  
  # combine all results in a df
  lapply(c(1:9), function(x) {get(paste0("m", x))}) %>%
    bind_rows() %>%
    mutate(dataset = x)
}) %>%
  bind_rows()  

```

* The low HR w/ the old dataset in 2013 looks like it is driven by 2018, which was the last data year for the old dataset. It is also when incidence is artificially high, which may or may not matter here (highest incidence is associated with very low air pollution levels so higher air pollution looks "protective"?).    
* This effect is not there with the new PM2.5 ST model

note that an end year of 2017 only cuts off a few person-years from 2018 in the old dataset

```{r}
models_old_new_time %>%
  filter(grepl("^ST_|SP_|^MM_", covariate)) %>% 

  ggplot(aes(y=description, x=hr, xmin=lower95, xmax=upper95, col=dataset)) +
  facet_wrap(~pollutant_predictors) +
  geom_vline(xintercept = 1, alpha=0.5, linetype=2) +
  geom_pointrange(position = position_dodge2(width = 0.3)) + 
   labs(title = "Dementia incidence HRs for 10 yr Exposures (no other pollutants in models)",
         x= "HR", y="Person-Years", col = "ST PM2.5 Version")
 
```

**why is the old model different from what Rachel reported?**    
* the "old" PM2.5 model dataset came from my files, not Rachel's (?)   
* the new models include exposures for the full 2018 year (more person-years)
*



```{r}

# show all year/main model results
models_old_new_time %>%
  filter(grepl("^ST_|SP_|^MM_", covariate),
         grepl("All Years", description)
         ) %>%
  select(dataset, description, hr, lower95, upper95) %>%
  arrange(desc(description)) %>%
  kable(caption = "Comparison of HRs estimates using old & latest ST PM2.5 exposure models. Using 10 yr exposure. adjusting for median household income (vs NDI)", 
        digits = 2) %>% 
  kable_styling()

```

looking more specifically at models that are restricted to 2 yr calendar bins

* note: Rachel grouped 2016-2018, whereas we grouped 2016-2017 & 2018-2020. the 2016 analysis thus includes a little over 2 yrs from the old and only 2 yrs for the new dataset

```{r, results="hide"}
#yrs <- comp$cal_2yr %>% unique() %>% as.character() %>% as.numeric() %>% sort()
# why does 2018 has issues?
#yrs <- seq(1994, 2016, by=2)
yrs <- seq(1994, 2018, by=1)

# x = 2018
# d = "old"

models_yr_specific <- lapply(yrs, function(x) {
  lapply(unique(comp$dataset), function(d){
    comp %>%
      filter(dataset==d,
             #cal_2yr==x
             exposure_year == x
             ) %>%
      mutate(cal_2yr = droplevels(cal_2yr)) %>% 
      run_cox(.,pollutant_predictors = "ST_pm25", event_indicator = "dementia_now",
              other_predictors = c(#"cal_2yr", 
                                   "degree", "income_cat", "male", "race_white")) %>%
      mutate(description = paste0(x),
             dataset=d)
    })
  }) %>%
  bind_rows()

```

```{r, fig.height=8}
models_yr_specific %>%
  filter(grepl("^ST_|SP_|^MM_", covariate)) %>% 

  ggplot(aes(y=description, x=hr, xmin=lower95, xmax=upper95, col=dataset)) +
  facet_wrap(~pollutant_predictors) +
  geom_vline(xintercept = 1, alpha=0.5, linetype=2) +
  geom_pointrange(position = position_dodge2(width = 0.5)) + 
   labs(title = "Dementia incidence HRs for 10 yr Exposures (no other pollutants in models)",
         x= "HR", y="Year Included in Analysis", col = "ST PM2.5 Version")

```


# ANALYSES WITH NEW MODELS ONLY

```{r}
## Exposure Coverage Threshold

# --> 

```


## Exposures Over Time

Comparison of assigned 10 yr participant exposures using the ST PM2.5 vs 2019 SP PM2.5

* **these results make a temporal pattern less clear**

```{r, fig.width=8}
sur %>%
  select(study_id, exposure_year, SP_pm25, ST_pm25) %>%
  drop_na() %>%
  pivot_longer(cols = c("SP_pm25", "ST_pm25")) %>%
  mutate(exposure_year = factor(exposure_year)) %>%
  
  ggplot(aes(x=exposure_year, y=value, col=name)) + 
  geom_boxplot() + 
  labs(y="10 yr Participant Exposure", col="PM2.5 Model")
  
sur %>%
  select(study_id, exposure_year, SP_pm25, ST_pm25) %>%
  drop_na() %>%
  pivot_longer(cols = c("SP_pm25", "ST_pm25")) %>%
  #mutate(exposure_year = factor(exposure_year)) %>%
  group_by(#exposure_year, 
           name) %>%
  summarize(
    n=n(),
    min=min(value),
    mean=mean(value),
    max = max(value),
    SD = sd(value),
    IQR = IQR(value)
  ) %>%
  kable(caption = "Distribution of 10 yr participant exposures for the study period", digits = 2) %>%
  kable_styling()
  
```

## Cohort Characteristics Over Time

The table below is for person-years (not persons). In more recent years, person-years included in the analysis are:    

* a bit older    
* more educated     
* living in slightly less disadvantaged neighborhoods (based on 2000 NSES Census values)   
* more likely to be diagnosed with dementia   
* have lower ST PM2.5 exposures   

year groups have roughly similar person-years, with 2015+ having the least (see table above)

```{r}
t1_labels <- list(
  labels = list(
    male = "Sex",
    #intakeage = "Age at Intake",
    age_start_exposure = "Age at Baseline",
    follow_up_years = "Enrollment Duration",
    corrected_anydementia = "All-Cause Dementia Cases",
    dementia_now ="All-Cause Dementia Incidence",
    ad_nincds = "Alzheimer's Disease Dementia Cases",
    vad_dementia = "Vascular Dementia",
    mixed_dementia = "Mixed Dementia",
    other_dementia = "Other Dementia",
    non_ad_dementia = "Non-Alzheimer's Disease Dementia Cases",
    #birth_cohort = "Birth Cohort",
    race_white = "Race",
    degree = "Degree",
    nses_z_cx = "Neighborhood Disadvantage Index (NDI)",
    apoe = "APOE4 Genotype",
    apoe_available = "Included in Main Analysis",
    model_wt = "IPW for Modeling Cohort",
    above_avg_ufp_10_42 = "Above Avg (PNC)",
    pollutant_prediction_ufp_10_42 = "PNC (pt/cm3) Exposure",
    pollutant_prediction_no2 = "NO2 (ppb) Exposure",
    pollutant_prediction_bc = "BC (ng/m3) Exposure" #,  pollutant_prediction_pm25 = "PM2.5 (ug/m3)"
    ),
  units = list(
    #intakeage = "Years"
    age_start_exposure = "Years"),
  categoricals = list(
    male = list('1' = "Male",
                '0' = "Female"),
    corrected_anydementia = list('1' = "Yes",
                                 '0' = "No"),
    dementia_now = list('1' = "Yes",
                                 '0' = "No"),
    vad_dementia = list('1' = "Yes", 
                        '0' = "No"),
    mixed_dementia = list('1' = "Yes", 
                        '0' = "No"),
    other_dementia = list('1' = "Yes", 
                        '0' = "No"),
    non_ad_dementia = list('1' = "Yes", 
                        '0' = "No"),
    ad_nincds = list('1' = "Yes",
                     '0' = "No"
                                 ),
    race_white = list('1' = "White",
                      '0' = "People of Color"),
    degree = list('0' = "None",
                  '1' = "GED/High School",
                  '3' = "Bachelor's",
                  '4' = "Master's",
                  '5' = "Doctorate",
                  '6' = "Other"),
    apoe = list('1' = "Carriers",
                '0' = "Non-Carriers"),
    apoe_available = list('TRUE' = "Included in Main Analysis",
                          'FALSE' = "Not Included in Main Analysis"),
    above_average_ufp_10_42 = list('FALSE' = "PNC Below Avg",
                                   'TRUE' = "PNC Above Avg")
    )#, groups=list("", "PNC Conc", "")
  )

#t1_labels <- readRDS(file.path("Data", "Output", "table1_labels.rda"))

print("cohort characteristics, grouped by calendar years to include similar person-years")
sur %>%
  drop_na(apoe, ST_pm25) %>%
  mutate(py_range = cut_number(exposure_year, 4)) %>%  
  t1read(., t1_labels) %>%
  table1(~age_start_exposure +  male + race_white +degree + nses_z_cx + #income_cat +
           apoe + model_wt + dementia_now + SP_pm25 + ST_pm25 | py_range, 
         data=.)

# separate last few years  

print("cohort characteristics, highlighting last study years")
sur %>%
  drop_na(apoe, ST_pm25) %>%
  mutate(#py_range = cut_number(exposure_year, 4)
    period = ifelse(exposure_year <= 2016, "1994-2016",
                    ifelse(exposure_year == 2017, "2017",
                     ifelse(exposure_year ==2018, "2018",
                            ifelse(exposure_year ==2019, "2019",
                                   ifelse(exposure_year ==2020, "2020", NA)))))
         ) %>%  
  t1read(., t1_labels) %>%
  table1(~age_start_exposure +  male + race_white +degree + nses_z_cx + #income_cat +
           apoe + model_wt + dementia_now + SP_pm25 + ST_pm25 | period, 
         data=.)

```
Incidence is a rough estimate (person-years is based on the number of rows, not age_end_exposure-age_start_exposure, so true incidence is slightly higher b/c of a smaller denominator)


## Survival Models

### Time-stratified

The results using the new ST model show:    

* a slight temporal effect for ST PM2.5     
* as expected, SP PM2.5 generally tracks ST PM2.5 well although it is least similar during early years where the CI is much wider and most similar during more recent years    

* MM NO2 (TRAP) produces similar results as the ST NO2

**what is going on w/ SP PM2.5 in 1994-2000? cohort is different? **

```{r, results="hide"}

############## RUN MODELS ############## 
#current dataset
pollutants <- c("ST_pm25", "SP_pm25", "MM_pm25",
                "ST_no2", "SP_no2", "MM_no2")

models_time <- lapply(pollutants, function(x) {
  
  m1 <- sur %>%
  run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
  mutate(description = "All Years (1994-2020)")

  m2 <- sur %>%
    filter(exposure_year <= g1_end) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
    mutate(description = paste0("1994-", g1_end))
  
  m3 <- sur %>%
    filter(exposure_year >= g2_start & exposure_year <=g2_end) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
    mutate(description = paste0(g2_start, "-", g2_end))
  
  m4 <- sur %>%
    filter(exposure_year >= g3_start & exposure_year <=g3_end) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
    mutate(description = paste0(g3_start, "-", g3_end))
  
  m5 <- sur %>%
    filter(exposure_year >= g4_start) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
    mutate(description = paste0(g4_start, "-2020"))
  
  #others
  m6 <- sur %>%
    filter(exposure_year <=2018) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
    mutate(description = paste0("Stop early: 1994-2018"))
  
  m7 <- sur %>%
    filter(exposure_year <=2017) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
    mutate(description = paste0("Stop early: 1994-2017"))

  #pm5[1,]
  
  # combine all results in a df
  lapply(c(1:7), function(x) {
    get(paste0("m", x))
    }) %>%
    bind_rows() %>%
    mutate(variable = x)

}) %>%
  bind_rows() %>%
  mutate(variable = factor(variable, levels=pollutants),
         pollutant_predictors = factor(str_to_upper(pollutant_predictors), levels = c("PM25", "NO2")),
         exposure_model = factor(exposure_model, levels = c("ST", "SP", "MM"))
  )

```

```{r}
############## HR PLOTS ##############
models_time %>%
  filter(grepl("^ST_|SP_|^MM_", covariate)) %>%

  ggplot(aes(y=description, x=hr, xmin=lower95, xmax=upper95, col=exposure_model)) +
  facet_wrap(~pollutant_predictors) +
  geom_vline(xintercept = 1, alpha=0.5, linetype=2) +
  geom_pointrange(position = position_dodge(width = 0.3)) + 
   labs(title = "HRs for 10 yr Exposures (no other pollutants in models)",
         x= "HR", y="Person-Years Included", col = "Exposure Model", shape="")

```




```{r}
### Exposure response curves

# --> ? can I do this w/ time-varying exposures?
# https://shariq-mohammed.github.io/files/cbsa2019/1-intro-to-survival.html

```



### Low concentration times earlier on

analysis w/ subset of low concentration areas during the earlier years. Models are stratified by whether exposures were above/below the overall mean for the selected study years. 

**note: this analysis is strange b/c some ppl could end up in the high & low categories in different years**

```{r, results="hide"}
pollutant_list2 <- c("ST_pm25")

models_low_conc <- lapply(pollutant_list2, function(x) {
  m1 <- sur %>%
    filter(exposure_year <=2007) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    filter(get(x) < mean(get(x))) %>%
      run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
      mutate(description = "1994-2007, low PM2.5")
  
  m2 <- sur %>%
    filter(exposure_year <=2007) %>%
    mutate(cal_2yr = droplevels(cal_2yr)) %>% 
    filter(get(x) >= mean(get(x))) %>%
      run_cox(.,pollutant_predictors = x, event_indicator = "dementia_now") %>%
      mutate(description = "1994-2007, high PM2.5")
  
  # m6[1,]
  # combine all results in a df
  lapply(c(1:2), function(x) {get(paste0("m", x))}) %>%
    bind_rows() %>%
    mutate(variable = x)
}) %>%
  bind_rows()


############## HR PLOTS ##############

```

```{r}
models_low_conc %>%
  filter(grepl("^ST_|SP_|^MM_", covariate)) %>%

  ggplot(aes(y=description, x=hr, xmin=lower95, xmax=upper95, col=exposure_model)) +
  facet_wrap(~pollutant_predictors) +
  geom_vline(xintercept = 1, alpha=0.5, linetype=2) +
  geom_pointrange(position = position_dodge(width = 0.3)) + 
   labs(title = "HRs for 10 yr Exposures (no other pollutants in models)",
         x= "HR", y="Exposure Concentration", col = "Exposure Model", shape="")

```




# Changes in the PM2.5 composition (speciation) over time

Beacon hill has speciation data but:   
* it starts ~2000  
* there are a lot of measures to look at   

```{r, eval=F}
# annual data report: https://pscleanair.gov/DocumentCenter/View/4548/Air-Quality-Data-Summary-2020?bidId= 
# data: https://aqs.epa.gov/aqsweb/airdata/download_files.html

######################################################################
# READ IN SITE DAILY AVERAGES & CALCULATE ANNUAL AVERAGES
######################################################################

# --> HOW TO SELECT & SUMMARIZE CHANGES IN THE SPECIES OF INTEREST?

#no speciation readings in 1994 in King County

parameters <- read.csv(file.path("Data", "epa", "parameter_classes.csv")) %>%
  pull(Parameter.Code)


sp2019_0 <- read.csv(file.path("Data", "epa", 
                               #"daily_SPEC_2000.csv"
                               "daily_SPEC_2019.csv"
                               )) 
  
sp2019 <- sp2019_0 %>%  
  filter(State.Name=="Washington",
         County.Name == "King",
         
         # don't need everything
         # WRONG?
         #Parameter.Code %in% parameters 
         ) %>% distinct(Local.Site.Name) %>% View()
  group_by(Latitude, Longitude, Parameter.Name, Sample.Duration, Units.of.Measure, Local.Site.Name, 
           #ignoring this b/c produces multiple readings per site. they seem to be similar across duplicates? 
           # could select a primary method in the future
           #Method.Name 
           ) %>%
  summarize(mean = mean(Arithmetic.Mean))


unique(sp2019$Local.Site.Name)

```







